---
title: "Homework-25"
author: "Christopher Augustine"
format: html
editor: visual
---

setting things up 
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

reading the csv in 
```{r}
dgrowth<- read.csv(here("data", "Daphniagrowth.csv"))
head(dgrowth)
```

making a box and whisker plot looking at parasites and growthrate and introducing function coord_flip which just flips axis 
```{r}
ggplot(dgrowth, aes(x= parasite, y= growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip()
```
creating the table for later to use in the anova  
```{r}
model_grow<- lm(growth.rate ~ parasite, data= dgrowth)
```

looking at the diagnostic graphs from autoplot 
```{r}
autoplot(model_grow, smooth.color= NA) 
```
doing the anova test to see if there is an effect on the treatments 
```{r}
anova(model_grow)
```

getting a summary table to look at coefficients
```{r}
summary(model_grow)
```
getting the mean growth rates for the different treatment groups 
```{r}
sumDat<- dgrowth %>% 
  group_by(parasite) %>% 
  summarise(meanGR = mean(growth.rate))
View(sumDat)
```

