---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

```{r}
df_No_NA <- df[complete.cases(df), ]
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

There is sufficient evidence to reject the null hypothesis of that there is no significant difference between means of treatments in regards to growth rate (F=3.69, df=2,46, p=0.033) 

Why are there 2 DF for the Treatment effect?

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this:

```{r}
sum(!is.na(df$Growth_rate_mm))
```

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

49 - 3 = 46 = degrees of freedom for the residuals.

```{r}
summary(grow.mod)
```

Practice your interpretation:

There is a signficiant effect of treatment on growth rate (df=2,46, p= 0.032) Specifically treatment 2 is significantly different from the control (p=0.011) while treatment 1 does not have a significant difference (p= 0.081) 

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

No, adding 1 dose of fertilizer does not change growth rate significantly enough from the control. Though it does reach margin effect for p<0.1 

Does adding 2 doses of fertilizer change the growth rate significantly from the Control?

Yes, adding 2 doses of fertilizer increases growth rate significantly compared to control. 

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes (Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

What do you think?
Im seeing a maybe a significant difference between control and the treatment groups but nothing between the treatments looks significant enough. 

#### Create a model:

```{r}
grow.mod.no.NA <-
  lm(Growth_rate_mm ~ Treatment, data = df_No_NA) 
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA?

all looks fine but normality has a little deviation at the top end. 

#### Interpret results:

```{r}
anova(grow.mod.no.NA) 
```



Practice your interpretation.

Why are there 2 DF for the Treatment effect?

3 levels, number of groups -1 =2 

Why are there 33 DF for the Residuals?

36 variables - 3 = 33 


How would you write a results statement?

We fail to reject the null hypothesis of no effect of which treatment on Growth Rate (F=2.791,2,33,p=0.076) 

Practice your interpretation:

```{r}
summary(grow.mod.no.NA)
```


Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

No adding 1 does of fertilizer does not have a significant effect on growth rate from the control.


Does adding 2 doses of fertilizer change the growth rate significantly from the Control?

Yes adding 2 doses of fertilizer does have a significant effect on growth rate from the control.

We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(data= df_All_NA, aes( x= Growth_rate_mm, y=Dried_mass_g))+
  geom_point(color="blue", size=2, alpha= 0.6)+
  labs(
    x= "Growth Rate (mm)",
    y= "Dried Mass (g)"
  )
  theme_bw()
```
Interpretation: 
seeing maybe a small positive relationship 
#### Create model

```{r}
regres_model1<- lm(Dried_mass_g ~ Growth_rate_mm, data = df_All_NA)
```

#### Check assumptions

```{r}
autoplot(regres_model1)
```

Don't worry if assumptions don't look perfect - focus on normality of the data.

#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(regres_model1)
```

```{r}
summary(regres_model1)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals?

number of groups 2, -1 = 1

2 options for the math we have 49 observations with 15 observations being excluded due to dried mass (49-15=34) or 45 observations with 11 being excluded due to grow rate (45-11=34).

Statement:
We can reject the null hypothesis that there is no relationship between dried mass and growth rate, as Growth rate increases so does dried mass (F=10.81, df=1,34, p=0.0236)

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(data= df_No_NA, aes( x= Growth_rate_mm, y=Dried_mass_g))+
  geom_point(color="blue", size=2, alpha= 0.6)+
  labs(
    x= "Growth Rate (mm)",
    y= "Dried Mass (g)"
  )
  theme_bw()
```

#### Create model

```{r}
regres_model2<-lm (Dried_mass_g ~ Growth_rate_mm, data= df_No_NA)
```

#### Check assumptions

```{r}
autoplot(regres_model2)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis?

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(regres_model2)
```
```{r}
summary(regres_model2)
```

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why?

again groups 2 then -1 = 1

2.  How many degrees of freedom are there for Residuals? Why?

observations(49)- regression coefficients (15)
49-15= 34 

3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis?

the data set if being used before or after omitting our NA values our degrees of freedom did not change so gonna guess R omitted them in the first model we did without omitted NA values. 