---
title: "Homework-26"
author: "Christopher Augustine"
format: html
editor: visual
---

setting everything up and reading in the data
```{r}
rm(list=ls())
library(tidyverse)
library(ggfortify)
library(here)
library(multcomp)
```

```{r}
pea<- read.csv(here("data", "s_r_pea_sections.csv"))
```

making treatment a factor
```{r}
pea$Treatment<- as.factor(pea$Treatment)
```

making a boxplot of the data to look at and see if I can make an assumptions by comparing treatments to treatments and to the control also flipping it to see the treatments
```{r}
ggplot(pea, 
  aes(x=Treatment, y=pea_section_length_ou))+
  geom_boxplot()+
  theme_minimal()+ coord_flip()
  
```
I am seeing a major difference between all treatments and the control, mainly between control and 2% glucose, fructose, and 1% 1% fructose glucose mix. 2% percent sucrose also seems to have a difference in mean but less then the other 3 stated above 

making a model to later check assumtions 
```{r}
model_pea<- lm(pea_section_length_ou ~ Treatment, data = pea)
View(model_pea)
```

plotting the model to check assumptions 
```{r}
autoplot(model_pea)
```

seems good only thing I notice is in Normal Q-Q for checking normaility of residuals most points are following a line with some deviations not much tho prob nothing


lets run the anova 
```{r}
anova(model_pea)
```

good small p value so we reject null hypo that there is no difference in means 

now for the tukeyhsd test to see differences between treatments 
```{r}
tukeyres<-glht(model_pea, linfct = mcp(Treatment="Tukey"))
```

summary to actually see that result
```{r}
summary(tukeyres)
```

getting letters display for easier interpitation 
```{r}
cld(tukeyres)
```
1% fruc + 1% gluc, 2% fruc, 2% gluco are not signifacntly different from eachother but are significalty different from 2% sucro and control 

Also control is significantly different from all other treatments as 2 % sucro is also significantly different from all other treatment groups 


