---
title: "Homework-23"
author: "Christopher Augustine"
format: html
editor: visual
---

getting things set up 
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

reading in and assigning the plant.growth.rate csv 
```{r}
plant_gr<- read.csv(here("data", "plant.growth.rate.csv"))
```

looking at the data 
```{r}
glimpse(plant_gr)
```
making a scatterplot of the data 
```{r}
ggplot(plant_gr,
       aes(x=soil.moisture.content, y=plant.growth.rate))+
  geom_point()+
  labs(
    y = "Plant Growth Rate (mm/week)") +
      theme_bw()
```
making a linear regression with the lm() function 
```{r}
model_pgr<- lm(plant.growth.rate ~ soil.moisture.content, data=plant_gr)
```

using the autoplot function which when provided a linear model by the lm() function will give 4 plots
```{r}
autoplot(model_pgr, smooth.colour = NA)
```

using the anova() function to make a sum of squares table
```{r}
anova(model_pgr)
```
using summary() function to make a table that reports model coefficient, standard errors, t and p values, R-squared values, adjusted R-squared, and F-statistics for the model overall
```{r}
summary(model_pgr)
```
slapping a linear model fitted line on our graph 
```{r}
ggplot(plant_gr,
       aes(x=soil.moisture.content, y=plant.growth.rate))+
  geom_point()+
  geom_smooth(method = lm)
  labs(
    y = "Plant Growth Rate (mm/week)") +
      theme_bw()
```

