---
title: "Homework-15"
author: "Christopher Augustine"
format: html
editor: visual
---

Homework 15 for biostat class looking into chapter 3 of thr gswr book. 
Date Oct-6-2024

setting things up 
```{r}
rm(list=ls())
library(tidyverse) 
library (here) 
library(dplyr) 
library(ggplot2)
```

importing the dataset "compensation.csv" and assigning it 
```{r}
compensation<- read.csv(here("data","compensation.csv"))
```

getting a summary of the data set and various statistics with the numerical data
```{r}
summary(compensation)
```
Using the function "select" to grab columns (adding a "-" in front of the column name gives all but that one)

```{r}
select(compensation, Fruit)
select(compensation, -Root)
```
using slice() to grab row 2 in the compensation dataset  
```{r}
slice(compensation, 2)
```
we can grab multiple rows by giving a set number of rows by using ":" in between row numbers 
```{r}
slice(compensation, 2:10)
```
also can grab specific rows by using c() 
```{r}
slice(compensation, c(2,5,12))
```
using filter to find rows were the statement given is true or false and to return it as a dataset (using with() gives them as a true or false set). 
```{r}
filter(compensation, Fruit> 80) 
```
using other R logical or boolean to examine the data more and also assigning a value to those outputs
```{r}
filter(compensation, Fruit==80.31)
filter(compensation, Fruit != 35) 
filter(compensation, Fruit >= 74)
filter(compensation, Root >75 | Root <10)
filter(compensation, Fruit >90 & Root <2.3)
equalto80.31<- filter(compensation, Fruit==80.31)
```

using mutate() to designate a new column with outputs from another column 
```{r}
compensation<- mutate(compensation, logFruit=log(Fruit))
head(compensation)
```

using arrange() to set the data set in a perticular order 
```{r}
arrange(compensation, Fruit)
```
using filter() and select() to get specifics from a column based on the vector answer from another and if its true or false

```{r}
select(filter(compensation, Fruit>80), Root) 
```

you can also use "pipe" 
```{r}
compensation %>% filter(Fruit > 80) %>% select(Root)
```

now using summarise() and group_by() and some math function and assigning it 
```{r}
fruit.mean <- summarise(group_by(compensation, Grazing), meanFruit=mean(Fruit))
```

again we can also use pipe and extend the math functions to do more
```{r}
compensation %>% group_by(Grazing) %>% summarise(meanFruit=mean(Fruit))
compensation %>% group_by(Grazing) %>% summarise(meanFruit=mean(Fruit), sdFruit=sd(Fruit))
```

