---
title: "Homework-16"
author: "Christopher Augustine"
format: html
editor: visual
---
This is a quarto document for biostat homework 16 code. 

clearing R's brain and installing packages to be used 
```{r}
rm(list=ls())
library(tidyverse)
library(here) 
```
reading the csv into R and assigning it in R's environment 
```{r}
cover<-read.csv(here("data","donnerville_percent_cover.csv"))
```

looking at the dataset 
```{r}
glimpse(cover)
head(cover)
```
making Site, Plot type, and cover type into factor variables 

```{r}
cover$Site<- as.factor(cover$Site)
cover$Plot.type<- as.factor(cover$Plot.type)
cover$cover_type<- as.factor(cover$cover_type) 
```

making the date a date variable 
```{r}
install.packages("lubridate")
library(lubridate)
```
```{r}
cover$Date <- dmy(cover$Date)
```

fixing errors in data
```{r}
levels(cover$Site)
fix1<- which(cover$Site%in% c("Donnerville 3","donnerville-3"))
cover$Site[fix1] <- "Donnerville 3"
 


fix2<- which(cover$Site%in% c("Donnerville 2", "donnerville 2"))
cover$Site[fix2] <- "Donnerville 2"


fix3 <- which(cover$Site%in% c("DONNERVILLE - 1", "donnerville 1", "Donnerville 1"))
cover$Site[fix3] <- "Donnerville 1"
cover$site<- factor(cover$Site)

fix4<- which(cover$cover_type%in% c("bare" , "Bare"))
cover$cover_type[fix4]<- "Bare"

fix5<- which(cover$cover_type%in% c("bryophtes", "Bryophytes"))
cover$cover_type[fix5]<- "Bryophytes"

fix6 <- which(cover$cover_type%in% c("vasc ","Vascular"))
cover$cover_type[fix6] <- "Vascular"
cover$site<- factor(cover$cover_type)
levels(cover$cover_type)
```
Determining average percent cover by type (bare, bryophytes, vascular plants) only in Donnerville 1 
```{r}
donnerville_1 <- cover %>% filter(Site == "Donnerville 1")
average_cover_donnerville1 <- donnerville_1 %>%
group_by(cover_type) %>%
summarise(Percent = mean(Percent))
```

Determining the average vascular%plant cover over all sites
```{r}

filter(cover, cover_type == "Vascular") %>% 
summarise(AveragePercentCover = mean(site))
```

looking at the frames
```{r}
average_cover_donnerville1
average_vascular_all_sites

```

