---
title: "Homework-18"
author: "Christopher Augustine"
format: html
editor: visual
---

Quarto document for homework 18 biostatistics class, looking into chapter 4 of the gswr book focusing on plots how exciting

getting things set up (tidyverse had ggplot2 and dyplr in it)

```{r}
rm(list=ls()) 
library(tidyverse)
library(here) 
```

reading the compensation.csv into the R environment and assigning it

```{r}
comp<- read.csv(here("data", "compensation.csv"))
```

looking at the data

```{r}
glimpse(comp)
```

using ggplot2 to make a simple scatterplot

```{r}
ggplot(comp, aes(x=Root, y=Fruit))+
  geom_point()
#aes is a function that defines the graphs mapping
# x and y equal to a column maps that to the corresponding axis 
```

everything in ggplot is built in layers so we always start with the type of code chunk above creating the baseline of the graph

adding a bw theme to the graph to get rid of the annoying white background

```{r}
ggplot(comp, aes(x=Root, y=Fruit))+
  geom_point() +
  theme_bw() 
```

increasing the size of the points using a size argument in the geom_point function

```{r}
ggplot(comp, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) +
  theme_bw() 
```

again always end the layers with theme_bw() but xlab and ylab ("xlabel","ylabel") push out the column and add the label given in QUOTES

```{r}
ggplot(comp, aes(x=Root, y=Fruit)) +
  geom_point(size = 5) +
  xlab("Root Biomass") + 
  ylab("Fruit Production") + 
  theme_bw() 
```

adjust the colours (weird how it being spelled different then my usual makes me remeber it easier)

```{r}
ggplot(comp, aes(x=Root, y=Fruit, colour = Grazing)) +
  geom_point(size = 5) +
  xlab("Root Biomass") + 
  ylab("Fruit Production") + 
  theme_bw() 
```

changing the shape of the points to correspond with the treatment

```{r}
ggplot(comp, aes(x=Root, y=Fruit, colour = Grazing, shape= Grazing)) +
  geom_point(size = 5) +
  xlab("Root Biomass") + 
  ylab("Fruit Production") + 
  theme_bw() 
```

box-and-whisker plots

```{r}
ggplot(comp, aes(x= Grazing, y= Fruit)) +
  geom_boxplot() +
  geom_point(size = 4, colour = "lightgrey", alpha = 0.5) + #size and colour we went over but alpha is transparency and that is equal to a number)  
  xlab("Grazing treatment") + 
  ylab("Fruit Production") + 
  theme_bw() 
```

histograms using geom=histogram() only give x variable computer does y

```{r}
ggplot(comp, aes( x = Fruit)) +
  geom_histogram()# but just this makes a ugly one 
```

so we do this

```{r}
ggplot(comp, aes( x = Fruit)) +
  geom_histogram(bins = 10) 
ggplot(comp, aes(x = Fruit)) + 
  geom_histogram(binwidth = 15) #this makes it pretty
```

we can also split the histogram into by treatment side note love how they justify this command in the book

```{r}
ggplot(comp, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) + 
  facet_wrap(~Grazing)
```

saving the histogram

```{r}
ggsave("Homework-18-Histogram.png")
```
